Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-08-09 GNproof.tex
minor edits for readability and notation

\begin{proof}
We have shown that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$. Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show: 
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm is  so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

@@

In order to exhibit a dependent subset of vertices in $J$, we first define a sequence of sets of propagators by
\[Q_0 = \Prop(j_m), \qquad Q_{l+1} = \Prop(I_i(Q_l))\quad \forall l \geq 0.\]
It is easily verified that this is an increasing chain of sets: indeed, if $q \in Q_l$ then $I_i(q) \in I_i(Q_l)$, and since $I_i(q)$ is one of the vertices that supports $q$ we must have $q \in \Prop(I_i(Q_l)) = Q_{l+1}$. Since $\cP$ is finite, the sequence therefore stabilizes in finitely many steps when $Q_{l+1} = Q_l$. Let $Q$ be this set of propagators.

Our remaining goal is to show that $Q \subseteq \mathcal{P}_{m-1}$. This will imply the theorem as follows.  Once the goal is proven we have $I_i(Q)\subseteq J$.  Then $I_i(Q) \cup {j_m}$ is a dependent set contained in $J$ because it only supports the propagators in $Q$ but has $|Q|+1$ vertices. Thus $J$ is a dependent set, and the Grassmann necklace $(I_1,\dots,I_n)$ constructed from $W$ does indeed define the positroid $M(W)$.

We now proceed to prove $Q\subseteq \mathcal{P}_{m-1}$ by inductively proving $Q_l \subseteq \mathcal{P}_{m-1}$ for all $l \geq 0$.  If $Q_0 = \Prop(j_m) = \emptyset$, then $Q = Q_0$, and $I_i(Q) \cup {j_m}= j_m$, a single vertex supporting no propagators, and thus a dependent set. Otherwise, $Q_0 \subseteq \cP_{m-1}$ follows immediately from the first observation above that $\Prop(j_m) \subseteq \cP_{m-1}$

Now assume that $Q_0,\dots,Q_{l-1} \subseteq \cP_{m-1}$, and consider $Q_{l}$. If $Q_{l} \setminus Q_{l-1} = \emptyset$ then we are done, with $Q = Q_l$.  Otherwise, let $q \in Q_{l} \setminus Q_{l-1}$. By construction there is some propagator $p_r \in Q_{l-1}\setminus Q_{l-2}$ such that $p_r$ and $q$ are both supported on $I_i(p_r) = a_r$. If $q$ is clockwise of $p_r$ on vertex $a_r$ then we immediately have $q \in \cP_{m-1}$ as observed above.  It remains to consider the case that $q$ is counter-clockwise of $p_r$ at $a_r$.

Note that we cannot have $a_r+ 1 = j_m$ when $q$ is counterclockwise of $p_r$, since $q \not\in \Prop(j_m)$ by assumption. If the other end of $p_r$ (i.e. the end not supported on $a_r$) lies on an edge between $a_r$ and $j_m$ then since $q$ is inside $p_r$ with respect to $<_{a_r-1}$ we must have $V(q) \subseteq [a_r-1,j_m]$ and hence $q \in \cP_{m-1}$ by Lemma [no fourth vertex]. In particular, if $l=1$ then $p_r \in Q_{0} = \Prop(j_m)$ and we must be in this case.

>>>>

\begin{proof}
We have shown that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$. Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show: 
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm assigns each propagator to a vertex in the support of that propagator, so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

@@

In order to exhibit a dependent subset of vertices in $J$, we first define a sequence of sets of propagators by
\[Q_0 = \Prop(j_m), \qquad Q_{l+1} = \Prop(I_i(Q_l))\quad \forall l \geq 0.\]
It is easily verified that this is an increasing chain of sets: indeed, if $q \in Q_l$ then $I_i(q) \in I_i(Q_l)$, and since $I_i(q)$ is one of the vertices that supports $q$ we must have $q \in \Prop(I_i(Q_l)) = Q_{l+1}$. Since $\cP$ is finite, the sequence therefore stabilizes in finitely many steps when $Q_{l+1} = Q_l$. Let $Q = \cup_{l \geq 0} Q_l$ be this set of propagators.

We will show that $Q \subseteq \mathcal{P}_{m-1}$, which will imply the theorem as follows: if $Q \subseteq \cP_{m-1}$ then we have $I_i(Q)\subseteq J$, and hence $I_i(Q) \cup \{j_m\} \subseteq J$ as well. Since $I_i(Q) \cup \{j_m\}$ only supports the propagators in $Q$ but has $|Q|+1$ vertices, it is a dependent set. Thus $J$ is a dependent set, and the Grassmann necklace $(I_1,\dots,I_n)$ constructed from $W$ does indeed define the positroid $M(W)$.

We now proceed to prove $Q\subseteq \mathcal{P}_{m-1}$ by inductively proving $Q_l \subseteq \mathcal{P}_{m-1}$ for all $l \geq 0$.  If $Q_0 = \Prop(j_m) = \emptyset$ then $Q = \emptyset$, and $I_i(Q) \cup \{j_m\}= \{j_m\}$ is a single vertex supporting no propagators, and thus a dependent set. Otherwise, $Q_0 \subseteq \cP_{m-1}$ follows immediately from the first observation above that $\Prop(j_m) \subseteq \cP_{m-1}$

Now assume that $Q_0,\dots,Q_{l-1} \subseteq \cP_{m-1}$, and consider $Q_{l}$. If $Q_{l} \setminus Q_{l-1} = \emptyset$ then we are done, with $Q = Q_l$.  Otherwise, let $q \in Q_{l} \setminus Q_{l-1}$. By construction there is some propagator $p_r \in Q_{l-1}$ such that $p_r$ and $q$ are both supported on $I_i(p_r) = a_r$. If $q$ is clockwise of $p_r$ on vertex $a_r$ then we immediately have $q \in \cP_{m-1}$ as observed above.  It remains to consider the case that $q$ is counter-clockwise of $p_r$ at $a_r$.

Note that we cannot have $a_r+ 1 = j_m$ when $q$ is counterclockwise of $p_r$, since $q \not\in \Prop(j_m)$ by assumption. If the other end of $p_r$ (i.e. the end not supported on $a_r$) lies on an edge between $a_r$ and $j_m$ then since $q$ is inside $p_r$ with respect to $<_{a_r-1}$ we must have $V(q) \subseteq [a_r-1,j_m]$, and hence $q \in \cP_{m-1}$ by Lemma [no fourth vertex]. In particular, if $l=1$ then $p_r \in Q_{0} = \Prop(j_m)$ and we must be in this case.
